### 2024-05-07 main_thesis_file.tex
Updates from Overleaf

%THIS IS THE MAIN FILE THAT TELLS LATEX WHICH FILES TO INCLUDE IN YOUR THESIS.

%THIS TEMPLATE PASSED DOWN FROM COURTNEY DRESSING
%EDITED SLIGHTLY AND UPLOADED TO OVERLEAF BY BEN COOK (bacook17@gmail.com)

% use "plain" for committee to read, or for printing single-spaced in general
\documentclass[twoside]{hvdthesis}

% use "bound" for submitting to Harvard 
%\documentclass[bound]{hvdthesis}

% no more microfilm?
%\documentclass[microfilm]{hvdthesis}

% Put all required packages and setup in one file
\input{setup}

% Put all your useful macros, shortcuts, and commands in one file
\input{mycommands}

\author{Soumya Roy}
\title{Thesis Title Goes Here}
\department{}
\subject{}
\month{Month}
\year{Year}
\advisor{Professor Durgesh Tripathi}

\begin{document}

% include the dissertation acceptance form
\includepdf[pages={1}]{placeholder_dac.pdf}

\clearpage
\thispagestyle{empty}
\begin{center}
\vspace*{\fill}
Page left intentionally blank.
\vspace*{\fill}
\end{center}

\frontmatter

	%% Title page
\makecover
	%% Copyright
%\copyright
	%% Abstract
\input{abstract.tex}
%\input{abstract_short.tex}
	%% Table of Contents
{\singlespace
\tableofcontents
}
\newpage
\clearpage

	%% List of Figures
\addcontentsline{toc}{chapter}{List of Figures}
\listoffigures
\newpage
	%% List of Tables
\addcontentsline{toc}{chapter}{List of Tables}
\listoftables

\newpage

\thispagestyle{plain}
\addcontentsline{toc}{chapter}{List of Abbreviations}
\chaptermark{List of Abbreviations}
\vskip 0.5cm
{\centerline {\Large \bf List of Abbreviations}}
\vskip 0.5cm
\normalsize

\input{abbr.tex}

\newpage
	%% Acknowledgments
%\input{acknowledge.tex}
\thispagestyle{plain}
\addcontentsline{toc}{chapter}{Acknowledgments}
\chaptermark{Acknowledgments}

\vskip 0.5cm
{\centerline {\Large \bf Acknowledgments}}
\vskip 0.5cm
\normalsize
\input{acknowledge.tex}

%\hfill -- RCH
%\newpage

\clearpage
%	%% Dedication
%\input{dedication.tex}
\thispagestyle{plain}
\addcontentsline{toc}{chapter}{Dedication}

\vspace*{\fill}
{\centerline {\em Your thoughtful dedication goes here.}}
%\vskip 1.5cm
%{\centerline {\em }}
\vspace*{\fill}

%\clearpage

\mainmatter
\pagestyle{fancy}

%% Uncomment the next four lines, plus the \begin{bibunit} and \end{bibunit} 
%% to use bibunits.  Read bibunits.dvi for more documentation
%\bibliographyunit[\chapter]
%\bibliographystyle*{apj}
%\bibliography*{apj-jour,planets}
%\bibliographyunit

\chapter[Introduction]{Introduction}\label{c:intro}
\input{./introduction/intro}
\clearpage
%
\chapter{Existing Solar Observations and Techniques: A bridge towards Solar Ultraviolet Imaging Telescope}\label{c:chap2}
\chaptermark{Solar Observatories and Analysis Techniques}
\input{./chapter2/ms.tex}
\clearpage
%
\chapter{Intital preparatory analysis for SUIT}\label{c:chap3}
\chaptermark{SUIT preparatory analysis}
\begin{quote}
{\em ~~~~~~~This thesis chapter originally appeared in the literature as} \\
{authors,
{\em journal reference info}}
\end{quote}
\input{./chapter3/ms.tex}
\clearpage
%
\chapter{Forward modelling SUIT observations}\label{c:chap4}
\chaptermark{SUIT forward model}
\begin{quote}
{\em ~~~~~~~This thesis chapter originally appeared in the literature as} \\
{authors,
{\em journal reference info}}
\end{quote}
\input{./chapter4/ms.tex}
%
\chapter{Stellar calibration of SUIT}\label{c:chap5}
\chaptermark{SUIT stellar calibration}
\begin{quote}
{\em ~~~~~~~This thesis chapter originally appeared in the literature as} \\
{authors,
{\em journal reference info}}
\end{quote}
\input{./chapter5/ms.tex}
%
\chapter{Effects of Solar flares on the local plasma environment from Mg II observations}\label{c:chap6}
\chaptermark{Solar flares Mg II}
\begin{quote}
{ ~~~~~~~This thesis chapter originally appeared in the literature as} \\
{{\em The Evolution of the Ratio of \ion{Mg}{2} Intensities During Solar Flares, Roy, S., \& Tripathi, D. 2024, ApJ, 964, 106, doi:\href{https://iopscience.iop.org/article/10.3847/1538-4357/ad2a46}{10.3847/1538-4357/ad2a46}}.}
\end{quote}
\input{./chapter6/ms.tex}
\clearpage
%
\chapter{Estimating thermal energy of two Solar Flares}\label{c:chap7}
\chaptermark{Solar flare energy}
\begin{quote}
{\em ~~~~~~~This thesis chapter originally appeared in the literature as} \\
{authors,
{\em journal reference info}}
\end{quote}
\input{./chapter7/ms.tex}
\clearpage
%
\chapter{First flares observed from {\suit}}\label{c:chap8}
\chaptermark{Solar flare energy}
\begin{quote}
{\em ~~~~~~~This thesis chapter originally appeared in the literature as} \\
{authors,
{\em journal reference info}}
\end{quote}
\input{./chapter8/ms.tex}
\clearpage
%

% I changed  thebibliography in hvdthesis.cls, so that it generates
% a ToC entry, and is headed References instead of Bibliography. -HJC
\singlespace
\bibliographystyle{apj}
\bibliography{your_bib_file}

\end{document}