### 2024-07-27 chapter7/ms.tex
Updates from Overleaf

\justifying

%%----------------------------------------------------
\section{Introduction} \label{sec:intro}
%%----------------------------------------------------

The Solar Ultraviolet Imaging Telescope onboard {\it Aditya-L1} \citep[{\it Aditya-L1}/SUIT,][]{article,ghosh16,adityal1,suit_main} provides a targeted probe into the Chromosphere and Transition region. It provides continuous full-disk and Region of Interest (RoI) coverage of the Sun in eleven pass bands. The details of these bands are provided in chapter~\ref{c:chap3} Tab.~\ref{tab:science_filters}. The eight narrow bands provide coverage across the \ion{Mg}{2} k and h lines, \ion{Ca}{2} h line, the CN band, red and blue wing of the \ion{Mg}{2} window and parts of the NUV continuum. As alluded earlier, this provides unprecedented coverage of the Chromospheric and Transition region structure of solar flares.

The {\it Aditya-L1} was launched on September 2nd, 2023. The first light observations were made on December 5th, 2023 while the payload was still in cruise phase. For the majority of the cruise phase the payload only operated in the 2k synoptic mode, i.e. it only captured continuous (2k $\times$ 2k) observations in NB4 (\ion{Mg}{2} h line) with one minute cadence. The L1 insertion was carried out on January 6th, 2024. Following that the payload also started taking (4k $\times$ 4k) observations in all eleven science filters. Various components of the flare detection algorithm were progressively turned on. In this chapter we discuss some of the initial flare observations made by SUIT. The first flare triggered by the flare detection algorithm was a X6.3 flare from north-west corner of the disk on February 22nd, 2024, which was also observed by several other instruments {\it e.g.} {\it SDO}/AIA, {\it IRIS}, {\it SO}/STIX. During the cruise phase, SUIT also observed a X1 flare from the south-east limb of the Sun on December 31st, 2023. We only have observations in the NB4 (\ion{Mg}{2} h line) channel for this flare. But this provides us with an unprecedented view of post-flare loops and possibly reconnection slingshot in the \ion{Mg}{2} line. Subsequent portions of this chapter will be divided into two parts, where we discuss the observations of these two flares separately.

%%----------------------------------------------------
\section{X6.3 flare observed on Feb 22nd, 2024} \label{sec:feb_22nd}
%%----------------------------------------------------

NOAA AR 13590 was visible on the north-east of the Solar disk on February 22, 2024. The AR was around a large sunspot in a cluster of sunspots, accompanied by a complex magnetic field structure. The active region flared multiple times during the same day, including an X1.7 flare that peaked around $\sim$ 06:32 UT, an M4.8 flare that peaked around $\sim$ 20:46 UT and an X6.3 flare that peaked $\sim$ 22:34 UT . \suit~is equipped with an onboard flare detection algorithm. Once the flare detection algorithm flags a flare and localizes the position of the flare on the detector, the program sequence prioritizes reading a fixed smaller Region of Interest (RoI) around the location for fast, higher cadence observation (for further details, please refer to \cite{flare_det}). The X6.3 flare was one of the first flares to be localized by the on-board flare detection algorithm. \suit~did not observe the X1.7 flare because, during that time, the payload was off-pointed to verify the stellar calibration program sequences. The flares were also observed by {\it SDO}/AIA, {\it SO}/STIX. {\it IRIS} observed the eastern edge of the X6.3 flare ribbons in a small [66\arcsec,62\arcsec] field of view (FoV) with a 4 step raster and 15 s raster cadence.

We show the light curve arising from the whole RoI observation in comparison to {\it GOES} flux in Fig.~\ref{fig:flare_full}. In Fig.~\ref{fig:flare_full}.b, we plot the AIA 1600 {\AA} (dashed blue line) and AIA 1700 {\AA} (black dot-dashed line), in comparison to the {\it GOES} 1 {--} 8 {\AA} light curve (Solid red line) for both the flares. In Fig.~\ref{fig:flare_full}.b, we show the GOES 1 {--} 8 {\AA} light curve in comparison to the NB3 (\ion{Mg}{2} k 279.6 nm, black dotted), NB4 (\ion{Mg}{2} h 280.3 nm, green dot-dashed) and NB8 (\ion{Ca}{2} h 396.85 nm, magenta dashed) light curves. NB4 light curve is offset from NB3 by -0.1 for better visibility. The NB8 light curve exhibits lower contrast and does not show a sharp peak compared to NB3 and NB4. The other interesting trend is exhibited by the continuum channels NB5, NB6 and NB7 in Fig.~\ref{fig:flare_full}.c, as a conspicuous rise in the continuum intensity is seen after both flares. In Fig.~\ref{fig:flare_full}.d we plot the {\it GOES} 1 {--} 8 {\AA} light curve in comparison to the STIX hard (25 {--} 50 keV, black dashed) and soft (5 {--} 10 keV, green dotted) X-ray light curve. The hard X-ray observation from STIX peaks at a time similar to that of NB3, NB4, and NB8.

%%%%%%%%%
\begin{figure}[ht!]
    \centering
    \includegraphics[width=0.8\textwidth,trim={2.3cm 2.5cm 1cm 4.5cm},clip]{lc_full_suit_contour.pdf}
    \caption{Light curves from the whole RoI FoV of \suit~observations compared with AIA, {\it GOES} and STIX observation for the two flares.}
    \label{fig:flare_full}
\end{figure}
%%%%%%%%%

The X6.3 flare provides a good example of the response of the local plasma environment to the flare in the Near Ultraviolet (NUV) regime in 200 {--} 400 nm. The flare peaked around $\sim$ 22:34 UT in the {\it GOES} observation. Images from six narrow band (NB) channels of \suit~ are shown in Fig.~\ref{fig:flare_nb3_peak} top panel at around $\sim$ 22:28-22:29 UT. This is at the peak of the NB3 (\ion{Mg}{2} k 279.6 nm) channel, as observed by \suit. The 60\% peak intensity contour of the NB3 intensity is marked with the black line in all figures of Fig.~\ref{fig:flare_nb3_peak} top panel. From the figure, we also see a similar structure in the NB4 (\ion{Mg}{2} h 280.3 nm) and NB8 (\ion{Ca}{2} h 396.9 nm). No similar structure is observed in the other continuum channels of Fig.~\ref{fig:flare_nb3_peak} top panel.

In Fig.~\ref{fig:flare_nb3_peak} bottom panel, we show the flare observations by \suit~ at their respective peaks. Again, we see very similar structures in NB3, NB4, and NB8. NB3 and NB4 peak at almost the same time $\sim$ 22:29 UT. Although NB8 shows a similar structure, the peak intensity of NB8 is observed slightly later around $\sim$ 22:29:41 UT. The peak intensities of the other continuum NB channels are observed progressively later. The NB5 (Red wing of the Mg window) peaks around $\sim$ 22:32:41 UT. Faint signatures of flare brightening are observed in NB5, marked with red arrows in Fig.~\ref{fig:flare_nb3_peak} bottom panel. No such signatures are observed as clearly in NB6 and NB7. Both of these channels peak around $\sim$ 22:28 UT.

%%%%%%%%%
\begin{figure}[ht!]
    \centering
    \includegraphics[trim={0cm 0.65cm 0cm 0cm},clip,width=0.5\textwidth]{suit_roi_nb3_peak.pdf} \\
    \includegraphics[width=0.5\textwidth]{suit_roi_all_peak.pdf}
    \caption{\suit~observations of the flare from various NB filters. Top panel: during NB3 peak. Bottom panel: During the peak of individual bands.}
    \label{fig:flare_nb3_peak}
\end{figure}
%%%%%%%%%

In Fig.~\ref{fig:flare_lc_suit}, we plot the light curve of the event to compare the observations across various bands. The AIA and \suit~ light curves in Fig.~\ref{fig:flare_lc_suit} are calculated by adding the counts within the region of 60\% peak intensity contour of NB3 (shown in Fig.~\ref{fig:flare_nb3_peak} top panel), after co-aligning and registering the AIA and \suit~observations and normalizing them to the peak intensity. In Fig.~\ref{fig:flare_lc_suit}.a, we show the GOES 1 {--} 8 {\AA} light curve in comparison to AIA 1600 {\AA} and AIA 1700 {\AA}. The AIA 1600 and 1700 {\AA} light curve peaks around $\sim$ 5 minutes earlier than the {\it GOES} peak.

In Fig.~\ref{fig:flare_lc_suit}.b, we show the GOES 1 {--} 8 {\AA} light curve in comparison to the NB3, NB4 and NB8 light curves. All the NB light curves behave remarkably similarly. The vertical dotted black line across all the panels in Fig.~\ref{fig:flare_lc_suit} denotes the peak intensity in NB3. NB3, NB4, and NB8 peaks around $\sim$ 22:29 UT, which is very similar to AIA 1600 {\AA} and 1700 {\AA}. We also plot the {\it GONG}-H$\alpha$ light curve from the \suit~contour region. The {\it GONG} light curve also peaks at around $\sim$ 22:29 UT. Both NB8 and {\it GONG}-H$\alpha$ shows less contrast variation in the light curve than NB3 and NB4.

We show the GOES 1 {--} 8 {\AA} light curve in comparison to the NB5 (Red wing of the Mg lines, blue dashed), NB6 and NB7 continuum channels in Fig.~\ref{fig:flare_lc_suit}.c. NB5 shows signs of flare response, although much weaker than NB3, NB4 and NB8. NB5 peaks around $\sim$ 22:31 UT, about $\sim$ 3 minutes later than NB3. Similar traits were observed from the images also, as pointed out in Fig.~\ref{fig:flare_nb3_peak} bottom panel and the accompanying discussions. More interestingly, NB6 and NB7 do not show the hallmark sign of a flare light curve, i.e. gradual increase and decrease in the intensity. These bands exhibit a slow but steady rise in intensity after the flare. Finally, in Fig.~\ref{fig:flare_lc_suit}.d, we show the {\it GOES} 1 {--} 8 {\AA} light curve in comparison to the STIX hard and soft X-ray light curve. The hard X-ray peaks around a similar time around $\sim$ 22:29 UT, similar to NB3, NB4 and AIA 1600 and 1700 {\AA}.

%%%%%%%%%
\begin{figure}[ht!]
    \centering
    \includegraphics[width=0.8\textwidth,trim={2.3cm 2.5cm 0cm 4.5cm},clip]{lc_suit_contour.pdf}
    \caption{Light curves from the pixels within intensity contour picked from \suit~and co-aligned AIA observations. The NB4 light curve in the second panel is offset by -0.1 from NB3 for better visibility. The vertical dotted dark line marks the flare's peak in NB3 observation.}
    \label{fig:flare_lc_suit}
\end{figure}
%%%%%%%%%

%%%%%%% ############# %%%%%%%
\subsection{XSM-spectra}\label{sec:xsm}
%%%%%%% ############# %%%%%%%

The Solar X-ray monitor on {\it Chandrayan-2}\citep[{\it Chandrayan-2}/XSM,][]{xsm} provides sun as a star soft X-ray spectra in 1{--}15 keV with 1s cadence and significantly better spectral resolution (180 eV at 5.9 keV) compared to STIX (1 keV ar 5.9 keV). The high spectral and temporal resolution allows us to measure the change in elemental abundances over the duration of the flares. XSM observed both flares with coverage in both impulsive and decay phase. The existing studies with XSM has successfully studies the thermal and elemental abundance evolution of various flares \citep{mondal21,kkepa23,nama23}.

Both the M and X-class flares were observed by XSM. The XSM 1{--}8 {\AA} light curves are plotted in Fig.~\ref{fig:xsm-obs} top panel. The red shaded region marks the time window where the Be filter was inserted to prevent saturation. There are periodic gaps in the data due to periodic lunar occultation. We show the spectra obtained by XSM in Fig.~\ref{fig:xsm-obs} bottom panel, during various phases of the flare. The violet spectra at 22:33 UT is near the flare peak, and shows a sharp decrease below 3 keV due to the insertion of the Be filter to avoid saturation. Various Mg, Al, S and Si lines are visible in the pre-flare spectra. In stark contrast during impulsive and peak of the flare we see strong signatures of Ca, Ar and Fe. Due to the uncertainty of the spectra during the Be window observation in $<~3~\mathrm{keV}$ regime, we fit the spectra beyond 3 keV throughput this analysis. We fit the spectra with the XSPEC model {\it chisoth}, which uses a wide range of pre-calculated spectra to fit the observed spectra (For more details please refer to the appendix of \cite{mondal21}).

%%%%%%%%
\begin{figure}[ht!]
\centering
    \includegraphics[trim={0.5cm 3.3cm 0.8cm 4cm}, clip, width=0.76\textwidth]{xsm_lc.pdf} \\
    \includegraphics[trim={1.8cm 2.5cm 3cm 3cm},clip,width=0.75\textwidth]{xsm_spec.pdf}
    \caption{Top panel: XSM observation of the two flares. The red shaded region marks the window when the Be filter was inserted. Bottom panel: X-ray spectra during various phases of the flare. The violet spectra is during the soft X-ray peak of the flare and shows a sharp decrease in the intensity below 3 keV. This is due to the insertion of Be filter to avoid saturation.}
    \label{fig:xsm-obs}
\end{figure}
%%%%%%%%

%%%%%%% ############# %%%%%%%
\subsubsection{Fitting XSM spectra}\label{sec:xsm-fit}
%%%%%%% ############# %%%%%%%

The observed soft X-ray spectra from the flaring plasma can be modelled by an isothermal plasma characterized by various parameters {\it e.g.} temperature, emission measure and abundance of various elements. As alluded earlier, we use the `{\it chisoth}' model \citep{mondal21} to fit the observed spectra from XSM. The `{\it chisoth}' model uses CHIANTI atomic database \cite{chianti} to calculate spectra for individual elements over a large temperature grid and stores as tables. The model includes elements from H (Z=1) to Zn (Z=30). The spectra of individual elements are calculated over a wide temperature range (0.3 {--} 50 MK). These modelled spectra of various elemental abundance over a wide temperature range are loaded in XSPEC and added with varying weights to fit the observed spectra. The Total model spectrum is given by  $$I_{mod}(T)~=~EM\sum_{X}I_{X}(T)N_{X}$$`EM' is the volume emission measure, $N_{X}$ is the abundance of element X relative to H and $I_{X}(T)$ is the modelled spectrum of element X at temperature T. The fit is done in a recursive process by minimizing the chi-squared between the observed and modelled spectra.

We initially used an isothermal model to fit the spectra. One of the key observations we make is a presence of systematic residual around the Fe complex around $\sim$ 6.5 keV. This is illustrated in the top panel of Fig.~\ref{fig:xsm_fit} top panel, where there is an excess visible in the blue side of the Fe line complex at $\sim$ 6.7 keV. \cite{mithun22} could not explain this systematic excess with multithermal DEM distributions. One of the possible explanation of this excess flux is Fe fluorescence emission. There have been observations of Fe line fluorescence at 6.4 keV (Fe K$\alpha$) and 7.06 keV (Fe K$\beta$) \citep{neupert67,doscheck71,bai79,tanaka84,parmar84,phillips12} using high-resolution X-ray spectra from Bent Crystal Spectrometer on-board Solar Maximum Mission \citep[Bent/{\it SMM},][]{bent,smm} and Yokoh \citep{yokoh} mission. Previous studies have suggested that this emission arises from the excitation of low ionization state Fe in the Photosphere either via the X-ray from the flaring plasma \citep{bai79} or directly from the non-thermal electron beam \citep{phillips73}. The Fe K$\alpha$ fluorescence is usually dependent on the position on the solar disk \citep{parmar84}. The emergent Fe K$\alpha$ emission suffers significant absorption and scattering along the line of sight, which increases with increasing heliocentric angle, resulting in a decrease in the observed intensity. For our event, the flaring region is near the disk center making the Fe K$\alpha$ fluorescence a valid candidate for explaining the excess. We added a Gaussian component to the `{\it chisoth}' model to get better fit (see Fig.~\ref{fig:flare_obs} bottom panel).

We add a Gaussian line component to our fit at 6.4 keV to explore the possibility of the excess flux arising from Fe K$\alpha$ fluorescence. We find that this fits the observed spectra better than previous instances, as demonstrated in Fig.~\ref{fig:xsm_fit} bottom panel. The K$\alpha$ emission would arise from the X-ray emission from the flaring regions having energies $>$ 7.12 keV, the K edge of Fe, exciting the Fe atoms in the Photosphere. We show the intensity of the fitted Fe 6.4 keV component excess (blue solid line), in comparison to the fitted flux in 7.12 {--} 8.5 keV (green dot-dashed line) in Fig.~\ref{fig:fe_excess} top panel. For reference, {\it GOES} 1{--}8 {\AA} soft X-ray flux (red dotted line) and STIX 25 {--} 50 keV Hard X-ray flux (black dashed line) are overplotted. STIX Hard X-ray is a fair representative of the non-thermal electron flux deposited into the foot points. In the bottom panel, the light curve fitted Fe 6.4 keV excess Gaussian component is plotted with the lightcurve from the bright kernels marked with the two boxes in Fig.~\ref{fig:flare_nb3_peak} bottom panel. In both panels, the peak time of the Fe excess (blue solid line), STIX hard X-ray (black dashed line) and NB5 brightness from box 1 (dotted magenta line) are marked with vertical lines.

%%%%%%%%%%
\begin{figure}[ht!]
    \centering
    \includegraphics[trim={0.5cm 1cm 0.5cm 0.7cm}, clip, width=0.77\textwidth]{xsm_fit.pdf}
    \caption{XSM spectra in 3 {--} 8.5 keV binned between 22:27:20 {--} 22:27:25 UT. Top panel shows the fit with ``chisoth+chisoth" model. Bottom panel shows the same spectra fitted with ``chisoth+chisoth+gaussian" model.}
    \label{fig:xsm_fit}
\end{figure}
%%%%%%%%%%

%%%%%%%%%%
\begin{figure}[ht!]
\centering
    \includegraphics[trim={1.3cm 0.1cm 0.3cm 1.2cm}, clip, width=0.95\textwidth]{fe_excess_4.pdf}
    \caption{Fe excess emission from 6.4 keV(blue solid) in comparison to STIX 25-50 keV(black dashed), GOES 1 {--} 8 $\AA$ (dotted red) and XSM 7.12 {--} 8.5 keV (green dot-dashed) light curve.}
    \label{fig:fe_excess}
\end{figure}
%%%%%%%%%%

%%%%%%% ############# %%%%%%%
\subsection{Discussion}\label{sec:dis1}
%%%%%%% ############# %%%%%%%

This section reports the \suit~narrow band imaging of the first localized flare by the onboard flare detection algorithm. We report the observation in NB3 (\ion{Mg}{2} k 279.6 nm), NB4 (\ion{Mg}{2} h 280.3 nm), NB8 (\ion{Ca}{2} h 396.9 nm) and the continuum channels NB5 (Red wing of \ion{Mg}{2}), NB6 and NB7. For both the flares, the NB3, NB4 and NB8 peak around the same time as AIA 1600 and 1700 {\AA}. For the X6.3 flare, NB3, NB4 behave very similarly to NB8 within the NB3 intensity contour (see Fig.~\ref{fig:flare_lc_suit}.b). But the NB8 peak over the whole active region is much broader compared to the NB3 and NB4 light curves (see Fig.~\ref{fig:flare_full}). This implies that NB8 behaves differently in other parts of the active region.

The NB5 observes the continuum that is usually attributed as the Balmer continuum. There have been previous studies where Photospheric metal lines went into emission and affected the Balmer continuum \citep{heinzel14,kleint17}. The dominant contribution would still be the Balmer continuum. \cite{reetika21} attributed the brightening in the SJI 2832 {\AA} continuum for a mini flare to direct signature of electron beams. \cite{kowalski19} showed for one event, that the SJI 2832 {\AA} continuum enhancement and several Phototspheric absorption lines going into emission can be attributed to significant Photospheric heating. The entire 2832 {\AA} window of IRIS had several \ion{Fe}{2} and \ion{Cr}{2} lines which are usually observed as absorption lines, in emission. Curiously enough, this observation was also made in a bright umbral flare kernel, similar to the current event. As there was no {\it IRIS} scan of the umbral brightening visible in NB5, unfortunately, we can not comment on the spectral nature of the bright kernel.

We see an excess around the 6.5 keV Fe complex from the XSM observations which can be fitted with a single Gaussian. From Fig.~\ref{fig:fe_excess}, the Fe excess light curve (blue solid line) behaves very similar to the soft X-ray flux beyond the Fe K edge at 7.12 keV (green dot-dashed line). The excess also shows no correlation with the STIX 25 {--} 50 keV hard X-ray flux (black dashed line), illustrating no significant contribution from the non-thermal electron flux. This suggests that the excess flux seen around 6.4 keV Fe complex arises from the Fe fluorescence from the flaring X-ray. If we assume the penumbral brightening observed in this flare to be similar as observed by \cite{kowalski19}, the bright kernels observed in NB5 mainly arises from a plethora of \ion{Fe}{2} lines. The flaring X-ray beyond Fe K edge (7.12 keV) photoionizes the Fe in the Photosphere, giving rise to both \ion{Fe}{2} lines in the red wing of \ion{Mg}{2}, along with the observed Fe fluorescence by XSM.

We see a similar brightening in NB2, the blue wing of the Mg window. The light curve of the brightening of the blue wing of Mg is shown in Fig.~\ref{fig:fe_excess} third panel. The light curve shows peak at both hard X-ray peak and later on closer to Fe excess component peak. This possibly shows both Photospheric and Chromospheric components from the NB2. Further investigation and modelling is required to comment on the local plasma parameters that would produce the bright kernels observed in both NB5 and NB2, with the relative timing within themselves and also in comparison to the various energies in X-ray.

The other interesting observation is the rise in the continuum intensity, specifically in NB6 and NB7 as the flares happen. We can see a steady rise in the continuum intensity after the M and X class flare (see Fig.~\ref{fig:flare_full}). For the X6.3 flare, we do see some signature of the flare in NB5, although it peaks about $\sim$ 5 minutes later compared to NB3 and NB4. The photospheric nature of the NB5 continuum explains the 5-minute delay of the peak from NB3. We see the flare peak in sequential order of formation height NB3 and NB4 (\ion{Mg}{2}, 22:29 UT) and NB8 (\ion{Ca}{2}) $\Longrightarrow$ NB5 (Photospheric continuum, 22:32:41 UT). We are observing the increase in continuum intensity from 283.2 nm to 388 nm. The consistent increase in continuum intensity is happening across the Balmer jump ($\lambda$~=~364.5 nm).

%%----------------------------------------------------
\section{X1 flare observed on Dec 31st, 2023} \label{sec:dec_31st}
%%----------------------------------------------------