### 2024-08-14 pub.tex

\section*{Refereed Journal Publications}

%%%%%%%%%%%
\begin{itemize}
    \item Soumya Roy and Durgesh Tripathi, 2024, ApJ, 964, 106, doi: \href{https://iopscience.iop.org/article/10.3847/1538-4357/ad2a46}{10.3847/1538-4357/ad2a46}
\end{itemize}
%%%%%%%%%%%

\section*{Technical Documents}

%%%%%%%%%%%
\begin{itemize}
    \item Roy, S., et al., "Solar Ultraviolet Imaging Telescope (SUIT) Handbook", ISRO Aditya-L1 mission
    \item Roy, S., et al., "Estimation of Effects of Space-craft Jitter on SUIT Imaging", ISRO Aditya-L1 mission
\end{itemize}
%%%%%%%%%%%