### 2024-08-16 introduction/intro.tex
Updates from Overleaf

\justifying

The Sun is our nearest astrophysical object, which serves as a natural laboratory for various branches of Physics - atomic physics, spectroscopy, turbulence, magneto-hydrodynamics, stellar evolution, dynamo and magnetism and plays a key role in \sroy{driving} space and terrestrial weather. The Sun is also the primary source of light and energy on earth. So it is only natural that we try to understand the different physical processes involved in the energectics of the Sun, which directly affects life on Earth. In this chapter I briefly introduce the Sun, various layers of its atmosphere \sroy{and solar flares}. \sroy{Finally we provide a motivation for the thesis and a brief outline.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Sun}\label{Sun}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The Sun is a G2V star, with surface luminosity $3.86~\times~10^{26}$ W and an effective temperature of $\sim 5780~K$. The Sun is mostly made out of Hydrogen(92.1\%) and Helium(7.8\%) along with negligible quantities of heavier elements C, N, O, Mg, Si, Ne and Fe\sroy{, among others}. The study of the Sun can be divided into three components: the Solar interior, the Solar atmosphere and the Heliosphere. Fig.~\ref{fig_solar_int} shows a schematic diagram of various layers of the Sun. \sroy{Below we briefly describe these three regions.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The solar interior}\label{solar_int}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In the innermost zone, the Sun has a temperature of about 15 MK and a density of $1.6~\times~10^{5}~km.m^{-3}$ that fuses Hydrogen into Helium mainly by p-p cycle \citep{bethe38} and to some extent by CNO cycle \citep{bethe39}. Beyond the core we have the optically thick radiative zone, where the Hydrogen and Helium are completely ionized. The high energy $\gamma$-ray photons generated in the core, have very small mean free path in this region, as they collide numerous times eventually thermalizing \sroy{and} becoming visible photon. Beyond the radiative zone the degree of ionization of Hydrogen and Helium changes drastically giving rise to a sharp temperature gradient \sroy{leading to convective instability}. This is the convective zone where the energy from the inner layers of the Sun is transported adiabetically to the surface of the Sun. In between the radiative and the convective zone, there is \sroy{believed to be }a thin layer, roughly at $\sim~0.7~R_{\odot}$ known as the tacholine. \sroy{This region} is believed to play a major role in generating Sun's magnetic field via the Dynamo process \citep{corbard01}.

%% ############# %%
\begin{figure}[ht!]
    \centering
    \includegraphics[width = 0.8\linewidth]{Figures/solar_int.png}
    \caption{A schematic depiction of the different \sroy{regions} of the Sun, \sroy{including} some of the most prominent magnetic structures on the Sun e.g Sunspot, prominence, coronal hole. Credit : NASA/Goddard.}
    \label{fig_solar_int}
\end{figure}
%% ############# %%

The core and the radiative zone of the Sun rotates as a solid body, whereas the convective zone rotates differentially. The rotation period varies from $\sim~25~days$ to $\sim~35~days$ from the equator to the pole.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The solar atmosphere}\label{solar_atmos}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The layers outside the convection zone together, form the solar atmosphere \sroy{and is further divided into different regions. }The Photosphere, which is also known as the Sun's surface, the Chromosphere, the transition region and the Corona. While we can define geometrical height or depths to define these layers, it is much more useful to define them using the local optical depths of various spectral lines, which are directly used to observe the layers. This also gives us a idea about which spectral features originate mainly from which portions of the solar atmosphere. Fig.~\ref{fig_solar_atm} shows the variation of temperature and number density across various layers of the Solar atmosphere \sr{Ref. here}.

%% ############# %%
\begin{figure}[ht!]
    \centering
    \includegraphics[width = 0.8\linewidth]{Figures/solar_atm.png}
    \caption{The variation of electron temperature (solid line) and electron density ($\mathrm{N_{e}}$, dashed line) and density of neutral hydrogen atoms ($\mathrm{N_{H}}$, dot-dashed line) in the solar atmosphere as derived based on the 1-D model calculations by \sr{Ref. here}.}
    \label{fig_solar_atm}
\end{figure}
%% ############# %%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The Photosphere}\label{photosphere}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The Photosphere is defined to be the layer where $\tau_{\lambda~\sim~5000}~=~1$. This is in the green part of the visible spectrum and the Sun is opaque in visible beyond this layer, hence it is called to be the surface of the Sun. This layer is 400-600 km \sroy{thick} and has an effective temperature of 5780 K. The magnetic filed lines arising from the tacholine penetrate the Photosphere and creates a `carpet' over the whole region \citep{priest14}. As mentioned earlier, this is the Solar surface known as the `Quite Sun'(QS). This region exhibits an average magnetic flux density of \textbf{10 -- 50~G}. The QS surface is covered with cells of roughly four size- granule, meso granules, super granules and giant cells. Some regions exhibit much stronger magnetic flux density often associated with highly twisted magnetic field structures. These magnetic features are manifested as Sunspots, spicules, \sroy{faculae} etc. \sroy{The Sunspots harbour some of the strongest concentrated magnetic fields, reaching up to several kilogauss in strength. The other noteworthy solar magnetic structure apart from these are believed to be spatially unresolved flux tubes composing a network with field strengths $\sim$ {\it $O(10^{3} G)$}} \citep{solanki93,grossman96,rubio19}.

As the density changes drastically at the Photosphere compared to the convective zone, the thermalized photons from Sun's core start free streaming, as the mean free path also increases drastically and as a result perturbs the thermal equilibrium (TE). \sroy{Therefore} the thermodynamic quantities of the Photosphere \sroy{are defined under the assumption of} Local Thermal Equilibrium (LTE). As we move \sroy{higher} from the Photosphere the LTE conditions also start deviating \sroy{because the density keeps on decreasing steadily} \citep{philips08}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The Chromosphere}\label{chromosphere}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The Chromosphere is a highly non-uniform dynamic layer with a thickness of 1500~--~3000~km, with increasing temperature (upto $10^{4}$ K) and decreasing number density. As seen in the Fig.~\ref{fig_solar_atm}, the temperature of the Chromosphere saturates before the dramatic rise in the transition region. This saturation is attributed to a steady deposition of acoustic energy by creation of shock waves \citep{carlsson07}. The Chromosphere also exhibits a sharp gradient in the plasma $\beta$ factor, non-LTE conditions, dominance of wave motions \sr{Ref. here}.

%%%%%%%%%%%%%%%%
\begin{figure}[ht!]
    \centering
    \includegraphics[trim={1cm 3cm 2cm 5cm},clip,width=0.8\textwidth]{Figures/sji_images.pdf}
    \caption{IRIS SJI Slit-jaw images of NOAA AR 13521 on Dec 21st, 2023 in \ion{Mg}{2} 2796 {\AA}~(left panel) showing the upper Chromosphere. The right panel shows the same active region in 2832 {\AA}~continuum corresponding to upper photosphere.}
    \label{fig:sji_features}
\end{figure}
%%%%%%%%%%%%%%%%

In Fig.~\ref{fig:sji_features} we show the NOAA AR 13521 in IRIS SJI 2796 {\AA} (left panel) and 2832 {\AA} (right panel). The dark Sunspots and plage regions surrounding it are clearly visible in the 2832 {\AA} wavelength window. In the 2796 {\AA} window we see the plage along with light bridges more clearly.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The Transition Region}\label{transition-region}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The transition region above the Chromosphere is a thin, dynamic layer where we witness a dramatic increase in temperature by two orders of magnitude along with a similar decrease in electron density, which is demonstrated in Fig. \ref{fig_solar_atm}. In general the this layer is roughly $\sim$ 100 km in thickness, but that can change depending on various dynamic conditions of Chromosphere($\sim~10^{4}~K$) below, \sroy{and/}or Corona($\sim~10^{6}~K$) on top. The transition region is generally characterized by the steep change in the temperature, pressure gradients, drastic change in local optical depth and competition between gas pressure and magnetic pressure. The transition region also manifests the ARs as various magnetic features such as small scale brightening, jets, spicules, fibrils etc.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The Corona}\label{corona}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The outermost layer of the atmosphere is the Corona. The Corona is only visible in naked eyes during the total solar eclipse or via coronagraphs. The coronal continuum spectrum is produced due to Thomson scattering of photons from photosphere off coronal electrons. Corona also emits in various emission lines, e.g. Mg, Fe, Si, S, Ca, O etc. across a wide band of electromagnetic spectrum ranging from X-ray to visible wavelengths. Similar to the other layers, the corona exhibits a variety of structures such as ARs, coronal holes, bright points, filaments, flares. The flare arcades are most prominently visible in the Corona. \sroy{The Corona also exhibits the most intense eruptions in the Solar system known as flares, which can be accompanied by Coronal Mass Ejections (CMEs). The Corona maintains a steady temperature of 1-2 MK during ambient conditions, but can rise up to tens of MK during solar flares.}

%%%%%%%%%%%%%
\begin{figure}[ht!]
    \centering
    \includegraphics[width=\textwidth]{Figures/aia_features.pdf}
    \caption{Image of the prominence eruption and associated flare from NOAA AR 11562 on Aug 31st, 2012$\sim$19:50 UT. The prominence is visible in all of the AIA channels. We see the coronal holes in AIA 335 \AA~\textbf{(panel c)}. The flare loops of the associated event is visible in AIA 94 \AA~\textbf{(panel e)} and AIA 131 \AA~\textbf{(panel f)}. The Sunspots are visible in HMI continuum \textbf{panel (h)}. We show a co aligned map of AIA 304 \AA, HMI magnetogram and AIA 131 \AA~in \textbf{panel (i)}. The spatial association of the Sunspots with the hot flareloop plasma and the prominence eruption is clearly demostrated here. }
    \label{fig:aia-feature}
\end{figure}
%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Solar Flares}\label{sol_flr}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Solar flares are the most powerful magnetic events in the solar system. They are described as a sudden increase in brightness in localized areas on Sun. Within tens of minutes, they can release over $10^{32}$ erg of energy, which is emitted across the entire electromagnetic spectrum from radio to gamma rays. They can also launch high energetic particles into the interplanetary medium. Most of the flares occur in magnetic active regions, and the amount of flare energy released is comparable to the free energy stored in the magnetic system \citep{emslie12,ash17}. The term "flare" is generally used explicitly for the entire magnetically-driven event's electromagnetic radiation, as it is the most significant fraction of the total energy liberated. The total energy released varies from event to event. It is also known that larger events occur much less frequently than smaller events.

In Fig.~\ref{fig:sji_features} we show one of the most well known and well studied solar eruptions from Aug 31st, 2012. A giant prominence erupted from the south east limb. The erupted prominence material was visible in all AIA channels from 304 {\AA} (Chromospheric He II, see Fig.~\ref{fig:aia-feature}a) across all six Coronal channels (Fig.~\ref{fig:aia-feature}b {--} g). The Sun spots were also clearly visible in HMI magnetogram and continuum (see Fig.~\ref{fig:aia-feature}h). In Fig.~\ref{fig:aia-feature}i we show the co-aligned AIA 304 {\AA} and HMI magnetogram image. The Sunspots lie at the base of the erupted prominence, accompanied by the hot thermal plasma associated with the accompanying flare (visible in Fig.~\ref{fig:aia-feature}e {--} f). The novelty of such observation lies in the clear spatial connection across various layers of the Sun, connected via energy and momentum transported from one layer to the other.

%% ##################################################### %%
\subsection{Brief history of flare observation}\label{sol_flr_1}
%% ##################################################### %%

In September 1, 1859, R.C. Carrington and R. Hodgson \sroy{independently} observed the first flare in the continuum of white light\citep{carrington1859,hodgson1859}. \sroy{Such} localized brightening on the Sun have remained an enigma ever since. We have been observing the local flaring events across all wavelengths from both ground and space based observatories.  Shortly after the observation by Carrington and Hodgson, people started studying the Sun extensively in the H$\alpha$ line which is formed in the Chromosphere, and the reports of flaring events became more and more frequent and progressively more and more complex. No two events were similar, as there were variations observed in source of size, ejections of plasma along with shock waves driving into the interplanetary space. Advances of radio technology during the second world war \sroy{affirmed} detection of presence of non-thermal electrons in the solar corona, during military radar operations \citep{hey46}. Around the same time S.E. Forbrush noticed ground level cosmic ray enhancement during major solar flares. These discoveries eluded that the flaring events do not only involve the thermal plasma, but somehow connect with high energy particles and involves the corona.

In the 1950s we started observing the Sun in hard X-rays ($\ge~10~keV$) with rockets and balloons. \cite{peterson59} discovered the first hard X-ray emission during a flare in 1958. Later on, it was deduced from the observations of the enhancements observed in radio and hard X-rays that, the the ejected energetic particles may contain a substantial fraction of the initial energy released \citep{brown71}. The hard X-ray is created by the bremsstrahlung of the electrons colliding into dense material, resulting in a power-law energy distribution. The broadband radio emission from 1 to 100 GHz is created from gyrosynchrotron emission. Finally, observations in EUV and soft X-ray($\le$~10~keV) have shown that the energy released form the flare heats the plasma contained in the coronal loops to temperatures beyond 30 MK.

%% ################################################################# %%
\subsection{Neupert effect}\label{npt_eff}
%% ################################################################# %%

\cite{neupert68} observed a curious correlation that, the soft X-ray flux during the rise phase of the flare is proportional to the time integral of the centimeter radio flux since the start of the flare. The centimeter radio flux is emitted by relativistic electrons. So, later on the same correlation was found between the hard X-ray flux and soft X-ray flux and can be expressed as,

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation}\label{npt_eff_eq}
    F_{SXR}(t)~\propto~\int_{t_{0}}^{t}~F_{HXR}(t')~dt'
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This empirical relation is known as the ``Neupert effect''. \cite{neupert68} had already suggested that this may be due to a causal relationship between the thermal plasma and the energetic electrons. The logical explanation is: the soft X-ray mainly originates from a thermal plasma heated by the energy of the flaring event deposited by the flare accelerated electrons. It is important to mention that we know now that Eqn.\ref{npt_eff_eq} is only valid if \sroy{cooling by conduction is negligible and/or the role of ions in the flare energy deposition was minimal.}

%% ################################################################# %%
\subsection{``Standard'' model of solar flare}\label{sol_flr_std_mod}
%% ################################################################# %%

\sroy{Based on various multi-wavelength studies of solar flares, combined with theoretical modelling a `standard' model of solar flares has been advocated by \cite{carmichael64,sturrock66,hirayama74,kopp76}. The flares occur due to abrupt release of magnetic free energy which was previously stored in the coronal magnetic field via flux emergence and surface motions \citep{forbes06}. Majority of the strongest flares are eruptive in nature. The `standard' model attributes the release of flare energy to magnetic reconnection in corona, which happens due to the eruption and/or coronal mass ejection \citep{shibata95,lin_n_forbes00,moore01,priest_n_forbes02}. It is noteworthy that, while over the years there have been continuous attempts this is not a all compassing quantitative model. Rather it is an attempt to orderly unify multi-wavelength observation and modelling attempts into a structure. Naturally the model can not explain all observations from various flares. A schematic picture of the model is shown in Fig.~\ref{fig:std_mod}. Below we describe the saliant points of the `standard' model:}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[ht!]
    \centering
    \hspace{5cm}
    \includegraphics[trim={1cm 8cm 0cm 8cm}, clip, width=0.95\linewidth]{Figures/std_mod.pdf}
    \caption{A schematic diagram of the standard model of solar flare.}
    \label{fig:std_mod}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%
\begin{enumerate}
    \item The magnetic reconnection happens in the corona, releasing the magnetic free energy. Electrons and energetic particles from the reconnetion site are accelerated along the realigned magnetic field lines. The accelerated particles that escape along the open field lines towards the earth give rise to the particle events seen from Earth. \sroy{The coronal reconnection source where the energy is released is often observed in HXR. The loop top HXR sources are often accompanied by the thermal SXR sources. The loop top HXR source is observed above the coronal SXR source \citep{masuda94}. Later on, most of the HXR sources were reported to be cospatial with the SXR source \citep{bataglia_n_benz07}. This thermal source is distinctly different from the post flare SXR and EUV loops. This thermal coronal source appears before the HXR emission begins, completely contrary to the "Neupert" behaviour. This SXR loop top source is a key behaviour of pre-flare phase. The emission measure of these sources are higher than typical coronal levels, suggesting evaporation from chromosphere before the HXR emission. This also suggests that during the early phases of flare, the energy is conducted gradually by thermal particles, compared to the energetic non-thermal particles later on.

    Given the existence of the thermal source, the coronal non-thermal source can be explained by a beam of non-thermal particles injected into the dense coronal thermal source. Some of the injected electrons escape the coronal source and precipitate lower. This model is known as the intermediate thin-thick target model (ITTT, proposed by \cite{wheatland95} and investigated further by \cite{metclaf99,fletcher98}.) While the ITTT model explains most of the observations from flares there are certain inconsistencies regarding the coronal X-ray source {\it e.g.} the source of coronal SXR in preflare conditions, difference in expected HXR brightness of the cornal source from models to observations still elude us. It is fair to summarise that even though we by and large understand under what ambient conditions how the coronal HXR source is created, we do not understand what underlying process cause such an environment to arise.}
    \item The precipitated particles move along the field lines downwards along the magnetic field lines, giving rise to the microwave \sroy{and radio} observations seen from flares via gyro magnetic radiation \sroy{at various frequencies dictated by the characteristics of local plasma}.
    \item The accelerated particles hit the Chromosphere, which is considerably denser than the Corona, giving rise to Hard X-ray and gamma-ray observed from the foot points. This also explains why the coronal hard X-ray source is considerably softer than the foot points. The energetic particles deposit their energy into the local chromosphere as they go through a series of collisions and eventually thermalize \citep{brown83}. \sroy{Foot point sources are occasionally detected on H-alpha flare ribbons that are approximately 30,000 km apart, marking the base of a magnetic arcade \citep{masuda01}.}

    \sroy{$\gamma$-ray emission is visible between 0.8 {--} 20 MeV generated by the accelerated ions. The neutron capture line at 2.2 MeV forming Deuterium is usually the strongest characteristic line visible from such emissions. Protons and other accelerated ions collide with the dense chromosphere and produce neutrons. These neutrons are captured by ambient protons to form Deuteron \citep{ramtay74,hua87}. There were later RHESSI observations \citep{hurford_2003,hurford_2006} that showed that the $\gamma$-ray footpoints did not coincide with the HXR footpoints. This demonstrated that the ions and electrons were accelerated differently, as proposed by \cite{emslie_2004}.}
    \item  As the plasma thermalizes, it emits in thermal bremsstrahlung, giving rise to the soft X-ray \sroy{radiation}. Also, as the energy is deposited into the Chromosphere from the accelerated particles, it heats the local Chromosphere environment it gradually increases the local pressure. As the pressure grows, when the pressure gradient builds up enough, the local plasma starts expanding upwards(essentially due to buoyancy) and slowly fills up the coronal loops with soft X-ray emitting plasma. This phenomenon is known as "Chromospheric evaporation". This was directly observed in blue-shifted lines of hot material. \sroy{This was first reported by \cite{doscheck80,feldman80}. \cite{antonucci89} reported velocities of 300 {--} 400 km/s with temperature $\sim$ 20 MK in \ion{Ca}{19}, while later on \cite{milligan06} reported much gentler up-flow with velocities $\sim$ 200 km/s in \ion{Fe}{19} line from {\it SOHO}/CDS observations. Several simulations also reproduced evaporation arising from non-thermal particle precipitation \citep{sterling93,hori98,reeves07}.}
    %\item This whole scenario explains the Neupert effect. As the energy form the accelerated particles is converted into the soft X-ray emitting plasma, and it builds up over time. That explains the soft X-ray being proportional to the time integral of the hard X-ray flux.
\end{enumerate}
%%%%%%%%%%%%%%%%%

This whole scenario is known as the ``Standard'' model of solar flare. While the standard model is an attempt to explain and unify various kind of differences seen from the numerous flares we observe, there are several cases where the standard model cannot explain the observations. For example, there have been observations of flares where the hard X-ray foot point do not form. It is proposed that in these cases the plasma in the coronal loops is so dense that the accelerated particle collide enough to thermalize within the flare loops before reaching the chromosphere, defusing the energy more evenly within the loop, rather than dumping it at the base of the loops \citep{veronig02,veronig04}. Another flare which previously occurred at the same region might explain the denser coronal loops \citep{strong84,bone07}.

%% ################################################################# %%
\subsection{The energectics of solar flares}\label{sol_flr_energ}
%% ################################################################# %%

With an experience of observing solar flares for more than 150 years, remarkably enough, we have barely started scratching the surface of the complexity involved with the solar flares. The re-configuring of magnetic structure, which almost always involves complex geometry, making almost all events unique in some sense. After that the released magnetic free energy is transported across various layers of the Sun and converted into various other forms of energy. \sroy{The overarching conversion of magnetic energy into various forms {\it i.e.} reconnection outflow, kinetic energy of particles, thermal and kinetic energy of chromospheric plasma is incredibly complex to model.}In case of the eruptive events, there is the added complexity of the kinetic and potential energy of the CMEs, the energy of the shocks and the kinetic energy of the solar energetic particles.
%Considering how the energy is transformed the magnetic energy of the active region that is released after the reconnection into the reconnection outflow jets, the kinetic energy of escaping particles, the thermal and the kinetic energy of the Chromospheric plasma evaporating, the radiative and conductive losses. In case of the eruptive events, there is the added complexity of the kinetic and potential energy of the CMEs, the energy of the shocks and the kinetic energy of the solar energetic particles.

In order to constrain the models of solar eruptions and various nuanced aspects of it, like magnetic reconnection, particle acceleration, heating etc. detailed quantitative characterization is absolutely necessary. There have been several studies that have tried to quantify the partition between various subsets of the energies. The questions that are particularly important are:

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{itemize}
    \item \textbf{If an active region can have enough free energy to account for the total energy released in the solar flares and/or CMEs.}
    \item \textbf{What is the energy partition between flares and CMEs.}
    \item \textbf{If the non-thermal component have enough energy to power up the thermal component.}
\end{itemize}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

It is generally well known by now that the active region have enough magnetic free energy to power the flare and CME \citep{emslie12,ash17}. The partition of energy between the flare and associated CME is much more fuzzy. \cite{emslie12} found that the flare and CME have energies of same order of magnitude, while \cite{ash17} concluded the flare dominates the CME in-terms of the energy. However the simple question, whether the non-thermal component of the flares have enough energy is still not resolved as even the most recent studies contradict each other in the most puzzling fashion. \cite{warmuth20} discussed the contradictions arising from some of the studies\citep{stosire07,emslie12,inglis14,warmuth16a,warmuth16b,ash17}. The details of the studies can be summarized as follows:

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[ht!]
    \centering
    \resizebox{\textwidth}{!}{%
    \begin{tabular}{||c|c|c|c|c|c|c||}
    \hline 
    \hline
       Study & No. of flares & GOES class range & Thermal model & Thermal spectrum & Thermal volume & Thermal losses \\
       \cite{stosire07} (S07 from hereon) & 18 & A3-B7 & Isotherm. & RHESSI & TRACE & X \\
       \cite{emslie12} (E12 from hereon) & 38 & C5-X28 & Isotherm. & RHESSI & RHESSI & Rad. \\
       \cite{inglis14} (IC14 from hereon) & 10 & B3-B9 & Multitherm. & RHESSI+AIA & RHESSI & Rad. \\
       \cite{warmuth16a,warmuth16b} (WM16 from here on) & 24 & C3-X17 & Isotherm. & RHESSI+GOES & RHESSI & Rad.,Cond. \\
       \cite{ash17} (A17 from here on) & 188 & M1-X7 & Multitherm. & AIA & AIA & Rad. \\
       \hline
    \end{tabular}}
    \caption{The details of the studies.}
    \label{tab1}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The bolometric energy serves as a representation of the overall energy released during solar flares. Regardless of how energy is released in a solar flare, whether through direct plasma heating, rapid bulk flows, or non thermal particle processes—the ultimate outcome is thermalization and subsequent radiation across the electromagnetic spectrum. Given that bolometric energy encompasses the entire spectrum, it is anticipated to reflect the total energy originally released. It's important to note that this specifically pertains to the energy liberated within the flare and does not encompass the energy associated with a concurrent coronal mass ejection (CME) or filament eruption. \cite{warmuth20} used the bolometric energy released in solar flares, to compare various studies on equal footing. Here we will discuss some of the main issues \cite{warmuth20} identifies comparing these studies.

The calculated thermal energy of all studies show an excellent correlation with the {\it GOES} class. Fig.~\ref{fig:goes-therm} left panel shows the peak thermal energy as a function of peak {\it GOES} flux for all the flares from the five studies. EM12 and WM16 are consistently one order of magnitude lower than the bolometric energy. Interpolation of E12 and WM16 are about half an order of magnitude lower compared to IC14, while S07 is very consistent with the extrapolation. The energies of A17 is about an order of magnitude higher compared to other studies, and even higher than the extrapolated bolometric energy for flares of same class from \cite{kretzschmar11} and E12. There can be several reason of these differences between the studies arising.

A possible reason can be attributed to different thermal models used to model the thermal component of the flares. {\it RHESSI} usually yields a higher temperature and lower emission measure compared to {\it GOES} due to the highly multi thermal nature of the flaring plasma \citep{bataglia05,ryan14,warmuth16a}. Although the {\it GOES} temperature is higher by a factor of 1.4, it is not dependant on the flare class. Along with that, \cite{warmuth20} also investigated whether assuming a isothermal or multi thermal model introduces any differences in the thermal energy estimates. They reported no significant differences due to differences of models (refer to Fig. 2 and the corresponding discussion in \cite{warmuth20}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[ht!]
    \centering
    \includegraphics[width=0.45\textwidth,trim={0cm 0cm 0cm 0.02cm},clip]{goes_flux_therm.jpg}
    \includegraphics[width=0.45\textwidth,trim={0cm 0cm 0cm 0.02cm},clip]{goes_flux_vol.jpg}
    \caption{{\it GOES} peak thermal energy (left panel) and volume of the thermal plasma (right panel) as a function of the peak {\it GOES} flux for all the flares from five studies (figure credit: \cite{warmuth20}). The C values show the linear correlation coefficient for various studies.}
    \label{fig:goes-therm}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The thermal energy estimated usually depends on the source volume, $U_{Th}\sim V$. Apart from S07 and A17, all other studies used {\it RHESSI} imaging to estimate the flare volume with a filling factor $f=1$. S07 determined the volume by employing a semicircular loop model, utilizing the cross-sectional area and loop length derived from the observed areas and separations of foot point brightening of 1600  {\AA} by {\it TRACE}. In contrast, A17 estimated the volume based on the flare area exceeding a certain threshold in the emission measure obtained from the spatial synthesis DEM method. Fig.~\ref{fig:goes-therm} right panel shows the volumes used for various studies to estimate the thermal energy as a function of the peak {\it GOES} flux. It is worth noting that the volume estimates from E12 and WM16 align closely, as do those from A+17, despite employing entirely different methodologies. In contrast, the micro flare volumes reported by IC14 are notably larger, ranging from one to two orders of magnitude beyond what would be anticipated based on the findings of the other four studies. Intriguingly, this discrepancy aligns with the additional volume required to account for the higher thermal energies observed in IC14. The uncertainty in detecting volume can be due to the CLEAN algorithm systematically overestimates the source size \citep{warmuth13a}. There has also been studies demonstrating that {\it RHESSI} has difficulties in resolving small sources \citep{dennis09,warmuth13b}, that would be applicable for small thermal sources in the micro flares. This illustrates the requirement for better volume estimation for better estimates of the thermal energy of flares.

We can conclude a few more puzzling questions from these studies as a whole:

%%%%%%%%%
\begin{itemize}
    \item The primary challenge in estimating the thermal energy of the hot plasma stems from determining the temperature distribution of the plasma using Differential Emission Measure (DEM), and the radiating plasma volume.
    \item The dissipation of energy from the hot plasma is substantial. Although there is widespread agreement regarding radiative losses across studies, the extent of conductive losses disagree significantly between the studies.
    \item The non-thermal energy in injected electrons is strongly dependent on the poorly constrained low-energy cutoff.
    \item The thermal and non-thermal energy partition changes with flare class. In smaller flares, there appears to be a deficiency of energetic electrons, whereas in larger flares, the injected non-thermal energy seems to sufficient in powering the thermal component. Does this signify the existence of an additional third heating source?
\end{itemize}
%%%%%%%%%

\sroy{In addition to these, there is one more looming question, related to the "standard" model of flares itself, that has been largely unanswered, not due to lack of efforts, rather limited mostly by the technological limitations dictating the detectability of the current instruments. In our discussion in \S\ref{sol_flr_std_mod}, we discussed about the fundamental differences inferred between the acceleration mechanisms of electrons and ions inferred from the different locations of the HXR and $\gamma$-ray foot points. The production of $\gamma$-ray from nuclear reactions in chromosphere opens up the possibility of high-energy nuclear reactions producing pions ($\pi^{0}$ and $\pi^{\pm}$) along with secondary neutrinos ($\nu$) through the decay of the pions. Such reactions would be driven by protons with energy $>$ 300 MeV \citep{ramtay75,hudson_n_ryan_95}. Detection of neutrinos will open up a new window in constraining the effects of the accelerated particles in the lower solar atmosphere, {\it i.e.} chromosphere and photosphere and the general characteristics of the local plasma environment. Unfortunately, in spite of considerable efforts no flare neutrinos have been detected so far. It is understood that the first hurdle is to detect a minute transient signal on top of the large solar neutrino background. This informs our required detection thresholds driven by the current ion acceleration models \citep{sol_nu_kamiokande,abe_2022}. If the inability to detect flare neutrinos persists, it indicates that not enough protons with energies $>$ 300 MeV is produced even in very large flares. That would impose an upper limit of the amount of energy that is deposited in particle acceleration from flaring events and/or the efficiency of the accleration mechanism.}

%% ################################################################# %%
\section{Motivation}\label{sec:mot}
%% ################################################################# %%

The main motivation of this thesis is to add understanding of effects of flares on the surrounding environment in lower and upper solar atmosphere, to answer some of the questions posed above. We have used imaging and spectroscopic observations from multiple space based observatories ({\it e.g.} AIA, HMI, IRIS, XRT, etc.). This research contributes to creating a perspective of flare energetics and its effect on the surrounding plasma environment from Chromosphere to Corona.

The solar flares are the strongest magnetic event in our solar system connecting various layers of the solar atmosphere ({\it e.g.} from Photosphere to Corona), coupled via mass and energy transport. While the Coronal manifestation of solar flares have been studied extensively with several instruments, there are only a few imaging and spectroscopic instruments observing the Photosphere to transition region. These include the {\it IRIS} \citep{iris}, AIA 1600, 1700, and 304  {\AA} \citep{aia} and some the filters of {\it Hinode}/SOT \citep{sot}. In addition to these, we also have had balloon experiment {\it SunRISE} \citep{Sunrise1,Sunrise2}, which provided high resolution imaging of the transition regiopn, from the SuFI instrument \citep{sufi}.

This illustrates the necessity of more observatories operating at the transition region wavelengths for more comprehensive understanding about how the solar flares manifest and affect the dynamics and heating of the lower solar atmosphere. In this regard, {\suit} onboard Aditya-L1 will play a crucial role in probing the lower solar atmosphere. The various imaging channels of {\suit} would observe the manifestation of flares in transition region at various heights (NB3 $\implies$ \ion{Mg}{2} k 279.6 nm, NB4 $\implies$ \ion{Mg}{2} h 280.3 nm, NB8 $\implies$ \ion{Ca}{2} h 396.8 nm), further lower in the transition region (NB2 $\implies$ blue wing of Mg window, NB5 $\implies$ red wing of Mg window) and Sunspots and possible signature of umbral brightening due to flares at the Photosphere (NB6 300 nm and NB7 388 nm) in full disk. This would also allow us to probe the interaction of the flaring region, with other active regions on disk, giving us a comprehensive overview of the effect of solar flares and their manifestation in the lower solar atmosphere.

%% ################################################################# %%
\section{Outline of Thesis}\label{sec:outline}
%% ################################################################# %%

The work within this thesis spans across various layers of the solar atmosphere to understand how solar flare couples these layers. It is carried out partly, as various preparatory analysis for {\suit}, forward modelling {\suit} observations, designing the stellar calibration scheme for {\suit}, analysis of existing transition region observations in anticipation of {\suit} observations, analysis of coronal observations of solar flares, and analysis of first solar flare observations by {\suit}. Here we will briefly discuss the structure of the rest of the thesis:

%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{itemize}
    \item In chapter \ref{c:chap2}, we review various space based missions available and used in this thesis. We also discuss various analysis techniques used in this thesis. We also motivate the importance of {\suit} and its main science goals. We also briefly discuss the specific questions regarding solar flares we can address with {\suit}.
    \item In chapter \ref{c:chap3}, we discussed some of the initial preparatory analysis we did for {\suit}, including the throughput model for {\suit}, how the filters mounted on {\suit} were chosen, the analysis of spacecraft jitter simulation and what effect it would have on the imaging. We also describe the stellar calibration scheme and the selection criteria for the standard star target. We use a Sun as a star spectra to evaluate the calibration scheme with the aforementioned throughput model.
    \item In chapter \ref{c:chap4}, we discuss the forward modelling pipeline developed for {\suit}. We use MPS-Atlas simulation and {\it IRIS} observations, with measured transmission profiles of various optical components and the PSF to forward model {\suit} observations. We also discuss a preliminary version of the deconvolution algorithm that will be deployed in the data reduction pipeline.
    \item In chapter \ref{c:chap5}, we use existing {\it IRIS} observations of three solar flares to probe how the lower solar atmosphere is affected by the flares. The ratio of the \ion{Mg}{2} k and h lines, is a proxy for the optical depth of the surrounding plasma. Hence, we use the observations of \ion{Mg}{2} lines along HMI magnetic field measurements to investigate the effects of magnetic field, if any. \cite{kerr15} conducted a similar study and found signatures of localised heating in the line intensity ratio. We found, that the line intensity ratio shows a correlated increase and decrease with the flare light curve. We argue this to be signature of plasma flow associated with flare, and a change in density of the local environment. This acts as a prelude of what is achievable with {\suit}. We can use a similar method to give full disk context of optical depth, with respect to eruptive events.
    \item We address the issues raised in \S\ref{sol_flr_std_mod} regarding the uncertainties of volume estimation for flaring plasma and its effects on the thermal energy estimation in chapter \ref{c:chap6}. We use observations from {\it SDO}/AIA, {\it STEREO-A}/EUVI, {\it SO}/STIX, {\it GOES}/SUVI to estimate the thermal energy of two solar flares. We propose a new  using existing tools, to use observations from different vantages ({\it e.g.} AIA and EUVI from different angles) to accurately calculate the volume of the flare arcade and estimate the thermal energy. We show that the accurate estimate as a function of time can have effects on the accurate estimation of thermal energy and the partition between the thermal and non-thermal component.
    \item In chapter \ref{c:chap7} we discuss the first flare observations of {\suit}. 
\end{itemize}
%%%%%%%%%%%%%%%%%%%%%%%%%%%