### 2025-04-11 pub.tex

\section*{Refereed Journal Publications}

%%%%%%%%%%%
\begin{itemize}
    \item ``Evolution of the Ratio of \ion{Mg}{2} Intensities during Solar Flares", {\bf Soumya Roy} and Durgesh Tripathi, 2024, ApJ, 964, 106, doi: \href{https://iopscience.iop.org/article/10.3847/1538-4357/ad2a46}{10.3847/1538-4357/ad2a46}\footnotemark[1]
    \item ``Near and Mid-ultraviolet Observations of X-6.3 Flare on 2024 February 22 Recorded by the Solar Ultraviolet Imaging Telescope on board Aditya-L1", {\bf Soumya Roy}, Durgesh, Tripathi, Sreejith, P, \& Ramaprakash, A. N., et. al., 2024, ApJL, 981, L19, doi:\href{https://iopscience.iop.org/article/10.3847/2041-8213/adb0be}{10.3847/2041-8213/adb0be}\footnotemark[1]
    \item ``Photometric Calibration \& Spectral Validation of the Solar Ultraviolet Imaging Telescope onboard Aditya-L1", Janmejoy Sarkar, {\bf Soumya Roy}, A. N. Ramaprakash,... 2024, J. Astron. Telesc. Instrum. Syst. 11(1), 014005 (2025), doi:\href{https://www.spiedigitallibrary.org/journals/Journal-of-Astronomical-Telescopes-Instruments-and-Systems/volume-11/issue-1/014005/Photometric-calibration-and-spectral-validation-of-the-Solar-Ultraviolet-Imaging/10.1117/1.JATIS.11.1.014005.short}{10.1117/1.JATIS.11.1.014005}\footnotemark[1]
    \item ``Science Filter Characterization of the Solar Ultraviolet Imaging Telescope (SUIT) on board Aditya-L1", Janmejoy Sarkar, Rushikesh Deogaonkar, Ravi Kesharwani, ... {\bf Soumya Roy},... 2024, Exp. Astron., 59, 3 (2025), doi: \href{https://doi.org/10.1007/s10686-024-09973-5}{10.1007/s10686-024-09973-5}
    \item ``Solar Ultraviolet Imaging Telescope on board Aditya-L1", Durgesh Tripathi, A. N. Ramaprakash, Sreejith P,... {\bf Soumya Roy},... 2025, Sol. Phys., 300, 30, \href{10.1007/s11207-025-02423-1}{10.1007/s11207-025-02423-1}\footnotemark[1]
    \item ``X-class flare on Dec 31, 2023, observed by the Solar Ultraviolet Imaging Telescope", {\bf Soumya Roy}, Durgesh Tripathi, Sreejith P, \& A. N. Ramaprakash, et. al., 2025, ApJL, 983, L6, \href{https://doi.org/10.3847/2041-8213/adc387}{10.3847/2041-8213/adc387}
    \item ``Test and Calibration of the Solar Ultraviolet
Imaging Telescope (SUIT) on board Aditya-L1", Janmejoy Sarkar, Nived V. N., {\bf Soumya Roy}, Rushikesh Deogaonkar et al., 2025, Accepted for publication in Sol. Phys.

\end{itemize}
%%%%%%%%%%%

\section*{Submitted Articles}

%%%%%%%%%%%
\begin{itemize}
    \item ``Evolution of thermal and cumulative non-thermal energy in solar flares", {\bf Soumya Roy}, Sophie Musset, Katharine K. Reeves, Durgesh Tripathi, Christopher S. Moore, 2024, Submitted to ApJ\footnotemark[1]{}\footnotetext{Part of this thesis}
    \item ``Forward Modeling Observations of the Solar Ultraviolet Imaging Telescope (SUIT) on board Aditya-L1", {\bf Soumya Roy}, L. S. Anusha, Durgesh Tripathi, et al. 2024, Submitted to ApJ\footnotemark[1]{}\footnotetext{Part of this thesis}
    %\item ``Critical Science Plan of the Solar Ultraviolet Imaging Telescope (SUIT) onboard Aditya-L1 mission", Durgesh Tripathi, A. N. Ramaprakash, ... {\bf Soumya Roy}, in Preparation
\end{itemize}
%%%%%%%%%%%

\section*{Technical Documents}

%%%%%%%%%%%
\begin{itemize}
    \item Roy, S., et al., "Solar Ultraviolet Imaging Telescope (SUIT) Handbook", ISRO Aditya-L1 mission
    \item Roy, S., et al., "Estimation of Effects of Space-craft Jitter on SUIT Imaging", ISRO Aditya-L1 mission\footnotemark[1]{}
\end{itemize}
%%%%%%%%%%%